A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2023-07-31 tidyindex.qmd
Update Intro paragraph 2

To construct an index, experts typically first define an underlying concept of interest to measure. This concept is latent and often without direct measure but of social and public importance. Relevant and available variables are then defined, collected, and combined as an index using statistical methods to approximate the concept. This construction process is often not straight forward and many decisions need to be made, whether explicitly or not. For example, decision on which variables to include may depend on the data availability and the choice in other similar indexes. Indexes constructed from linear combination of variables need to decide on the weight assigned to each variable. Some indexes have a spatial and/or temporal component and variables can be aggregated to different spatial resolutions and temporal scales for different monitoring purpose. These decisions can result in different index values and have different practical implication.

>>>>

To construct an index, experts typically start by defining a concept of interest that requires measurement. This concept often lacks a direct measurable attribute or can only be measure as a composite of various processes, yet it holds social and public significance. To create an index, once the underlying processes involved are identified, relevant and available variables are then defined, collected, and combined using statistical methods into an index that aims at measure the process of interest. The construction process is often not straightforward, and decisions need to be made, such as the selection of variables to be included, which might depend on data availability and the statistical definition of the index to be used, among others. For instance, the indexes constructed from linear combination of variables need to decide on the weight assigned to each variable. Some indexes have a spatial and/or temporal components, and variables can be aggregated to different spatial resolutions and temporal scales, leading to various indexes for different monitoring purposes. Hence, all these decisions can result in different index values and have different practical implications.
